---
title: "Introduction to the uno package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the uno package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # set TRUE once package is installed
)
library(uno)
```

## Overview

The **uno** package provides a complete Uno card game engine built in R. It
uses three OOP systems:

| System | Used for |
|--------|----------|
| R6     | `Card`, `Deck`, `Player`, `Game` — all mutable mid-game |
| S4     | `GameConfig` — strict slot validation before anything runs |
| S3     | `GameResult`, `Leaderboard` — value objects with custom `print()`/`summary()` |

A high-performance Rcpp simulation engine allows you to compare strategies
across thousands of games in seconds.

---

## 1. Core classes

### Card

```{r card}
library(uno)

# Number card
red5 <- Card$new("red", value = 5L, type = "number")
red5

# Action card — no value needed
skip <- Card$new("blue", type = "skip")
skip$label()      # "blue Skip"
skip$is_action()  # TRUE

# Wild — must start as "black"
wild <- Card$new("black", type = "wild")
wild$is_wild()    # TRUE

# Compatibility check
top  <- Card$new("red", value = 5L, type = "number")
blue5 <- Card$new("blue", value = 5L, type = "number")
blue5$matches(top, current_colour = "red")  # TRUE  (same value)
```

### Deck

```{r deck}
deck <- Deck$new()
deck$size()   # 108

# Draw a card
card <- deck$draw()
deck$size()   # 107

# Deal a hand
hand <- deck$draw_n(7L)
length(hand)  # 7
```

### Player

```{r player}
alice <- Player$new("Alice", strategy = "aggressive")
alice$receive_many(hand)
alice$hand_size()   # 7

# Which cards can Alice legally play?
top_card <- deck$draw()
legal <- alice$legal_indices(top_card, current_colour = "red")
```

### GameConfig (S4)

```{r config}
cfg <- methods::new("GameConfig",
  n_players  = 2L,
  hand_size  = 7L,
  stacking   = FALSE,
  force_play = TRUE,
  seed       = 42L
)
```

Validation is strict — out-of-range values throw an error:

```{r config-error}
# Throws: n_players must be in 2:4
methods::new("GameConfig", n_players = 5L)
```

---

## 2. Playing a game manually

```{r manual-game}
cfg <- methods::new("GameConfig",
  n_players = 2L, hand_size = 7L,
  stacking = FALSE, force_play = TRUE, seed = 1L
)

specs <- list(
  list(name = "Alice", strategy = "aggressive"),
  list(name = "Bob",   strategy = "random")
)

game <- Game$new(cfg, specs)
game$top_card()$label()    # e.g. "red 7"
game$active_colour         # e.g. "red"

# Advance the game automatically (bots only)
while (!game$is_over()) {
  game$auto_turn()
}

# Inspect the result
res <- game$result()
print(res)
```

---

## 3. GameResult and Leaderboard (S3)

```{r results}
# print() shows winner and standings
print(res)

# summary() returns a tibble
summary(res)

# Build a leaderboard over multiple games
lb <- new_leaderboard()

for (seed in 1:5) {
  cfg2 <- methods::new("GameConfig",
    n_players = 2L, hand_size = 7L, stacking = FALSE,
    force_play = TRUE, seed = as.integer(seed)
  )
  g <- Game$new(cfg2, specs)
  while (!g$is_over()) g$auto_turn()
  lb <- add_game_result(lb, g$result())
}

print(lb)
summary(lb)  # win rates per player
```

---

## 4. Simulation with Rcpp

The Rcpp engine (`simulate_games()`) runs thousands of games in seconds.
Available strategies: `"random"`, `"aggressive"`, `"colour_matcher"`.

```{r sim-basic}
results <- simulate_games(
  n          = 5000L,
  strategies = c("random", "aggressive", "colour_matcher"),
  stacking   = FALSE,
  seed       = 42L
)

head(results)
```

### Benchmark: Rcpp vs pure-R

```{r benchmark}
library(microbenchmark)

microbenchmark(
  Rcpp = simulate_games(1000L, c("random","random")),
  R    = simulate_games_r(100L, c("random","random")),   # 100 only — R is slow
  times = 3L
)
```

### Strategy comparison study

```{r study}
grid <- list(
  c("random","random"),
  c("aggressive","random"),
  c("colour_matcher","random"),
  c("random","aggressive","colour_matcher")
)

study <- run_simulation_study(grid, n_per_combo = 2000L)
print(study[, c("strategies","win_rate_random","win_rate_aggressive",
                "win_rate_colour_matcher")])
```

---

## 5. Shiny dashboard

```{r shiny}
run_app()
```

The dashboard has three tabs:

- **Play** — interactive Uno against 1–3 bots. Click cards to play, use Draw /
  Pass buttons, choose wild colours via a dropdown.
- **Simulate** — configure strategies and game count, run the Rcpp engine,
  view win-rate bar charts and turn-length histograms.
- **Leaderboard** — all Play-tab games are recorded automatically. Sortable
  table and win-rate chart.

---

## 6. House rules

`GameConfig` exposes two house-rule toggles:

| Slot | Default | Effect |
|------|---------|--------|
| `stacking` | `FALSE` | Allow a player to stack +2 on +2 (or +4 on +4) instead of drawing |
| `force_play` | `TRUE` | Player **must** play if a legal card exists (standard rule) |

```{r house-rules}
cfg_house <- methods::new("GameConfig",
  n_players = 3L, hand_size = 7L,
  stacking   = TRUE,   # stacking enabled
  force_play = TRUE,
  seed       = NA_integer_
)
```

---

## Package structure

```
uno/
├── R/
│   ├── 01-card.R          # Card R6 class
│   ├── 02-deck.R          # Deck R6 class
│   ├── 03-player.R        # Player R6 class
│   ├── 04-game-config.R   # GameConfig S4
│   ├── 05-game.R          # Game R6 orchestrator
│   ├── 06-results.R       # GameResult + Leaderboard S3
│   ├── 07-simulate-r.R    # Pure-R simulation (benchmark reference)
│   ├── 08-simulate-cpp.R  # Rcpp wrapper + study function
│   └── run_app.R          # run_app() launcher
├── src/
│   └── uno_sim.cpp        # Rcpp simulation engine
├── inst/app/
│   └── app.R              # Shiny dashboard
├── tests/testthat/
│   ├── setup.R
│   ├── test-card.R
│   ├── test-deck.R
│   ├── test-player.R
│   ├── test-game-config.R
│   ├── test-game.R
│   └── test-results.R
└── vignettes/
    └── intro.Rmd          # This document
```
